```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(car)
library(pander)
#library(agricolae)
#library(DescTools)
library(stats)
library(BSDA)
#library(pgirmess)
#library(multcomp)

### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.
# by Reid Ginoza

p.value.string = function(p.value){
  p.value <- round(p.value, digits=4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}

```



```{r}
redwine <- read.csv("winequality-red.csv", sep=";")
whitewine <- read.csv("winequality-white.csv", sep=";")
wine <- rbind(redwine, whitewine)
type <- as.factor(c(rep("Red Wine", nrow(redwine)), rep("White Wine", nrow(whitewine))))
wine <- cbind(wine, type)
```


## 1. One Way Anova Test on the contain of Sulfure Dioxide

```{r}
wine.Summary <-  wine %>% group_by(type) %>% summarise("total sulfur dioxide mean " = mean(total.sulfur.dioxide), " total sulfur dioxide stdev" = sd(total.sulfur.dioxide))
pander(wine.Summary, style='rmarkdown') 
```

```{r}
wine_anova_results <- aov(total.sulfur.dioxide ~ type, data = wine)
wine_anova_table <- summary(wine_anova_results)
pander(wine_anova_table, style='rmarkdown') 
```

#### Hypothesis test

**Hypotheses**

$H_0: \ \mu_1 = \mu_2$ <br> 
$H_1: \ \mu_1 \ne \mu_2$

**Test Statistic**

$F_0 = `r round(wine_anova_table[[1]][["F value"]][1],digits=2)`$.

***p*-value**

$`r p.value.string(wine_anova_table[[1]][["Pr(>F)"]][1])`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that the total sulfure dioxide contain in the two wine are diferents.


## 2. Kruskal-Wallis Test on the contain of Alcohol

```{r}
wine.Summary2 <-  wine %>% group_by(type) %>% summarise("alcohol median" = median(alcohol), "alcohol IQR" = IQR(alcohol))
pander(wine.Summary2, style='rmarkdown') 
```

```{r}
wine_KW_results <- kruskal.test(alcohol ~ type, data = wine)
pander(wine_KW_results, style='rmarkdown') 
```

**Hypotheses**

$H_0: \ \mbox{M}_1 = \mbox{M}_2$ <br> 
$H_1: \ \mbox{M}_1 \ne \mbox{M}_2$ 

**Test Statistic**

$H = `r round(wine_KW_results$statistic,digits=2)`$.

***p*-value**

$`r p.value.string(wine_KW_results$p.value)`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that there is a diferent in the alcohol contain of the two type of wines.


## 3. Variability Test of pH 

```{r}
pHvar <- var.test(redwine$pH, whitewine$pH, alternative = "t", conf.level = 0.95, ratio = 1)
pander(pHvar, style='rmarkdown') 
```

**Hypotheses**

$H_0: \sigma_r = \sigma_w$
$H_1: \sigma_r \ne \sigma_w$

**Test Statistic**

$F_0 = `r round(pHvar$statistic, digits = 2)`$.

***p*-value**

$`r p.value.string(pHvar$p.value)`$.

**Rejection Region**

Reject $H_0$ if $p < \alpha$, where $\alpha=0.05$. $`r p.value.string(pHvar$p.value)` \nless 0.05.$

**Conclusion and Interpretation**

Fail to reject $H_0$. There is insufficient evidence to suggest that there is a difference in the variability of pH between red and white wines.


## 4. Chloride Content Test

Formally test to determine if the mean of the chlorides contain on red wine is greater than that of white wine. Use the appropriate version of the t-test with the critical at the α = 0.05 level.

```{r}
wine.Summary <-  wine %>% group_by(type) %>% summarise("chlorides mean" = mean(chlorides), "chlorides stdev" = sd(chlorides))
pander(wine.Summary, style='rmarkdown') 
```

```{r}
chlorides_t <- t.test(redwine$chlorides, whitewine$chlorides, paired = FALSE, alternative = "greater")
chlorides_p <- p.value.string(chlorides_t$p.value)
pander(chlorides_t, style='rmarkdown')
```

**Hypotheses**

$H_0: \ \mu_R \le \mu_W$ <br> 
$H_1: \ \mu_R > \mu_w$

**Test Statistic**

$t_0 = `r round(chlorides_t$statistic, digits=3)`$.

***p*-value**

$`r chlorides_p`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that the contain chlorides in the red wine are greather than the white wine.


## 5. Variability Test of Residual Sugar 

```{r}
rsvar <- var.test(redwine$residual.sugar, whitewine$residual.sugar, alternative = "t", conf.level = 0.95, ratio = 1)
pander(rsvar, style='rmarkdown') 
```

**Hypotheses**

$H_0: \sigma_r = \sigma_w$
$H_1: \sigma_r \ne \sigma_w$

**Test Statistic**

$F_0 = `r round(rsvar$statistic, digits = 2)`$.

***p*-value**

$`r p.value.string(rsvar$p.value)`$.

**Rejection Region**

Reject $H_0$ if $p < \alpha$, where $\alpha=0.05$. $`r p.value.string(rsvar$p.value)` < 0.05.$

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that there is a difference in the variability of residual sugar contain between red and white wines.


## 6. Citrid Acid Content Test   

Formally test to determine if the mean of the citrid acid contain on red wine is less than that of white wine. Use the appropriate version of the t-test with the critical at the α = 0.05 level.

```{r}
wine.Summary3 <-  wine %>% group_by(type) %>% summarise("citric acid mean" = mean(citric.acid), "citric acid stdev" = sd(citric.acid))
pander(wine.Summary3, style='rmarkdown') 
```

```{r}
citricacid_t <- t.test(redwine$citric.acid, whitewine$citric.acid, paired = FALSE, alternative = "less")
citricacid_p <- p.value.string(chlorides_t$p.value)
pander(citricacid_t, style='rmarkdown') 
```

**Hypotheses**

$H_0: \ \mu_R \ge \mu_W$ <br> 
$H_1: \ \mu_R < \mu_w$

**Test Statistic**

$t_0 = `r round(citricacid_t$statistic, digits=3)`$.

***p*-value**

$`r citricacid_p`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that the contain of citric acid in the red wine is less than the white wine.


## 7. Proportion Test on Density

```{r}
wine_density1 <- z.test(wine$density, alternative = "greater", mu = 1, sigma.x = sd(wine$density))
pander(wine_density1, style='rmarkdown') 
```

**Hypotheses**

$H_0: \ \mu \le 1$ <br>
$H_1: \ \mu > 1$

**Test Statistic**

$z = `r wine_density1$statistic`$.

***p*-value**

$p = `r wine_density1$p.value`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that the mean density wine is greater than 1.

```{r}
wine_density2 <- z.test(wine$density, alternative = "less", mu = 0.98, sigma.x = sd(wine$density))
pander(wine_density2, style='rmarkdown') 
```

**Hypotheses**

$H_0: \ \mu \ge 0.98$ <br>
$H_1: \ \mu > 0.98$

**Test Statistic**

$z = `r wine_density2$statistic`$.

***p*-value**

$p = `r wine_density2$p.value`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that the mean density wine is less than 0.98.

```{r}
wine_density3 <- z.test(wine$density, alternative = "less", mu = 0.99, sigma.x = sd(wine$density))
pander(wine_density3, style='rmarkdown') 
```

**Hypotheses**

$H_0: \ \mu \ge 0.99$ <br>
$H_1: \ \mu > 0.99$

**Test Statistic**

$z = `r wine_density3$statistic`$.

***p*-value**

$p = `r wine_density3$p.value`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that the mean density wine is less than 0.99.


## 8. Regression model

```{r}
one_model <- lm(quality ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + total.sulfur.dioxide + density + pH + sulphates + alcohol, data=wine)
one_coef <- coefficients(one_model)
one_anova <- anova(one_model)
one_summary <- summary(one_model)
one_t <- as_tibble(one_summary[[4]])
one_ci <- as_tibble(confint(one_model, level=0.95))
```

The resulting regression model is 
\[ \hat{y} = `r one_coef[[1]]` + `r one_coef[[2]]`x_1 + `r one_coef[[3]]`x_2 + `r one_coef[[4]]`x_3 + `r one_coef[[5]]`x_4 + `r one_coef[[6]]`x_5 + `r one_coef[[7]]`x_6 + `r one_coef[[8]]`x_7 + `r one_coef[[9]]`x_8 + `r one_coef[[10]]`x_9\]

### ANOVA table and significance test of the regression line

The corresponding ANOVA table is as follows

```{r}
pander(one_anova, style='rmarkdown') 
```

**Hypotheses**

$H_0: \ \beta_1 = 0$ <br>
$H_1: \ \beta_1 \ne 0$

**Test Statistic**

$F_0 = `r round(one_anova$"F value"[1], digits=2)`$.

***p*-value**

$`r p.value.string(one_anova$"Pr(>F)"[3])`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Comparing every p value of the previous table with the alpha, can be check that with at lest one of then is greather than alpha, by this reason fail to reject $H_0$. There is not sufficient evidence to suggest that the regression line is significant.

### Significance test of the individual predictor

```{r}
pander(one_t, style='rmarkdown') 
```

**Hypotheses**

$H_0: \ \beta_1 = 0$ <br>
$H_1: \ \beta_1 \ne 0$

**Test Statistic**

$t_0 = `r round(one_t$"t value"[4], digits=2)`$.

***p*-value**

$`r p.value.string(one_t$"Pr(>|t|)"[4])`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Comparing every p value of the previous table with the alpha, can be check that with at lest one of then is greather than alpha, by this reason fail to reject $H_0$. There is not sufficient evidence to suggest that the regression line is significant.

